QPLIB: A Library of Quadratic Programming Instances

### 2017-01-28 QPLIB-3.tex
update pictures
- use most current numbers
- corrected count on nonconvex constraints
- use positive eigenvalues in Q^0 if maximization

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\centering
  \includegraphics[width=0.85\textwidth]{pic_overview.png}
  \caption{Distribution of variables and constrains  of the qplib
instances \label{fig:distribution}}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\centering
  \includegraphics[width=0.85\textwidth]{pic_var.png}
  \caption{Number of variables \label{fig:pic_var}}
\end{figure}

%\begin{figure}\centering
%  \includegraphics[width=0.85\textwidth]{pic_var_small.png}
%  \caption{Number of variables (a) \label{fig:pic_var_small}}
%\end{figure}
%
%\begin{figure}\centering
%  \includegraphics[width=0.85\textwidth]{pic_var_medium.png}
%  \caption{Number of variables (b) \label{fig:pic_var_medium}}
%\end{figure}
%
%\begin{figure}\centering
%  \includegraphics[width=0.85\textwidth]{pic_var_big.png}
%  \caption{Number of variables (c) \label{fig:pic_var_large}}
%\end{figure}

Similarly, Figures~\ref{fig:pic_constr} 
%Figure~\ref{fig:pic_constr_medium}
%and Figure~\ref{fig:pic_constr_big} 
describes how the number of linear and quadratic
constraints are distributed within the instances. The instances are sorted
accordingly to the total number of constraints.
% and divided in the same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
For each instance we report the total number of constraints with a ``$+$''
and the total number of quadratic constraints
with a ``$\times$''. Also, in this case, different percentages of linear and
quadratic constraints are present and well-distributed across the spectrum
of constraint sizes, although both medium- and large-size instances show a
prevalence of lower percentages of quadratic constraints. In particular,
from Figure~\ref{fig:pic_constr} we learn that while the maximum number
of linear constraints exceeds 100000, the maximum number of quadratic
constraints tops up at around 10000. This is, however, reasonable, considering
how quadratic constraints can, in general, be expected to be much more
computationally challenging than linear ones, especially if nonconvex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\centering
  \includegraphics[width=0.85\textwidth]{pic_constr.png}
  \caption{Number of constraints \label{fig:pic_constr}}
\end{figure}

%\begin{figure}\centering
%  \includegraphics[width=0.85\textwidth]{pic_constr_small.png}
%  \caption{Number of constraints (a) \label{fig:pic_constr_small}}
%\end{figure}
%
%\begin{figure}\centering
%  \includegraphics[width=0.85\textwidth]{pic_constr_medium.png}
%  \caption{Number of constraints (b) \label{fig:pic_constr_medium}}
%\end{figure}
%
%\begin{figure}\centering
%  \includegraphics[width=0.85\textwidth]{pic_constr_big.png}
%  \caption{Number of constraints (c) \label{fig:pic_constr_big}}
%\end{figure}

Figure~\ref{fig:pic_conv_constr} shows the instances with at least one quadratic constraint sorted according to the number of quadratic constraints (vertical axis). For each instance we report the total number of constraints with a ``$+$''
and the total number of nonconvex quadratic constraints
with a ``$\times$''.
It can be seen that the majority of instances only have convex constraints; of those that have nonconvex ones, however, a significant fraction have no convex ones at all.

\begin{figure}\centering
  \includegraphics[width=0.85\textwidth]{pic_quad_conv_vs_nonconv.png}
  \caption{Quadratic constraints (instances with at least one) \label{fig:pic_conv_constr}}
\end{figure}

Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} focuses on
the instances with a quadratic objective function, ordered by the
the percentage of negative eigenvalues in the $Q^0$ matrix (vertical axis). The instances are mostly
clustered around two values. About 25\% of the instances have convex objective function,
i.e., they have 0\% of negative eigenvalues. Among the others, a vast
majority has around 50\% of negative eigenvalues. However, instances
with high or low percentages of negative eigenvalues are present, too.

\begin{figure}\centering
  \includegraphics[width=0.85\textwidth]{pic_neg_eig.png}
  \caption{Negative eigenvalues (instances with quadratic objective function) \label{fig:pic_neg_eig}}
\end{figure}

Similarly, Figure~\ref{fig:pic_density} shows the instances with a
quadratic objective function sorted according to the density of the
$Q^0$ matrix (vertical axis). The majority of the instances have either
a very low or a rather high density: indeed, about 30\% of the instances
have density smaller than 5\%, and about 30\% of the instances have
density larger than 50\%. However, also intermediate values are present.

\begin{figure}\centering
  \includegraphics[width=0.85\textwidth]{pic_density.png}
  \caption{Density of $Q^0$ (instances with quadratic objective function) \label{fig:pic_density}}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\centering
  \includegraphics[width=0.55\textwidth]{pic_overview.pdf}
  \caption{Distribution of number of variables and constraints of QPLIB instances
instances. \label{fig:distribution}}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\centering
  \includegraphics[width=0.55\textwidth]{pic_var.pdf}
  \caption{Number of variables of QPLIB instances. \label{fig:pic_var}}
\end{figure}

Similarly, Figures~\ref{fig:pic_constr} (left)
%Figure~\ref{fig:pic_constr_medium}
%and Figure~\ref{fig:pic_constr_big} 
describes how the number of linear and quadratic
constraints are distributed within the instances. The instances are sorted
accordingly to the total number of constraints.
% and divided in the same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
For each instance we report the total number of constraints with a ``$+$''
and the total number of quadratic constraints
with a ``$\times$''. Also, in this case, different percentages of linear and
quadratic constraints are present and well-distributed across the spectrum
of constraint sizes, although both medium- and large-size instances show a
prevalence of lower percentages of quadratic constraints. In particular,
from Figure~\ref{fig:pic_constr} (left) we learn that while the maximum number
of linear constraints exceeds 100000, the maximum number of quadratic
constraints tops up at around 10000. This is, however, reasonable, considering
how quadratic constraints can, in general, be expected to be much more
computationally challenging than linear ones, especially if nonconvex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\centering
  \includegraphics[width=0.49\textwidth]{pic_constr.pdf}
  \includegraphics[width=0.49\textwidth]{pic_quad_conv_vs_nonconv.pdf}
  \caption{Number of constraints, quadratic constraints, and nonconvex quadratic constraints of QPLIB instances. \label{fig:pic_constr}}
\end{figure}

Figure~\ref{fig:pic_constr} (right) shows the instances with at least one quadratic constraint sorted according to the number of quadratic constraints (vertical axis). For each instance we report the total number of constraints with a ``$+$''
and the total number of nonconvex quadratic constraints
with a ``$\times$''.
It can be seen that the majority of instances only have nonconvex constraints. %; of those that have nonconvex ones, however, a significant fraction have no convex ones at all.

Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
the instances with a quadratic objective function, ordered by the
the percentage ``hard'' eigenvalues in the $Q^0$ matrix (vertical axis), by which
we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
i.e., they have 0\% of ``hard'' eigenvalues. Among the others, a vast
majority has around 50\% of ``hard'' eigenvalues. However, instances
with high or low percentages of ``hard'' eigenvalues are present, too.

\begin{figure}\centering
  \includegraphics[width=0.49\textwidth]{pic_neg_eig.pdf}
  \includegraphics[width=0.49\textwidth]{pic_density.pdf}
  \caption{``Hard'' eigenvalues (left) and density (right) of $Q^0$ matrix for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
\end{figure}

Similarly, Figure~\ref{fig:pic_neg_eig} (right) shows the instances with a
quadratic objective function sorted according to the density of the
$Q^0$ matrix (vertical axis). The majority of the instances have either
a very low or a rather high density: indeed, about 30\% of the instances
have density smaller than 5\%, and about 30\% of the instances have
density larger than 50\%. However, also intermediate values are present.
